### 2011-11-06 doc/pldi2012/circuit.tex
2

In this section we give an informal overview of Legion through an example.
Listing~\ref{lst:code_ex} shows Legion code for an electrical
circuit simulation, which illustrates the core features of the programming model.

The circuit simulation is an application that takes a collection of
circuit elements consisting of wires and nodes where wires meet.  
The simulation is a three-phase algorithm that repeatedly updates
currents, distributes charges, and updates voltages for as many
time steps as the simulation demands.  We first
describe the partitioning of regions for the simulation and how
the program exececutes in section \ref{subsec:partitioning}.  
We then describe the data type declarations 
and the constraints they enforce on the simulation that enable
the runtime to efficiently execute the application
in section \ref{subsec:datatypes}.

@@

Line 18 declares the main simulator function, which takes a
{\tt Circuit} to be simulated.  This function specifies the regions it
will access as well as the {\em permissions} and {\em coherence}
properties it requires of those regions.  In this case, the {\tt RWE}
annotation specifies that the function accesses the regions {\tt c.r\_all\_nodes}
and {\tt c.r\_all\_wires} with read-write permissions and exclusive
access (the strongest coherence mode).  The simulation of the circuit
will read and write all of the nodes and all of the wires, and it must
be done with exclusive access to ensure a correct result.  Functions that
declare their accessed regions, permissions, and coherence properties are
called {\em tasks} and will be considered by the Legion runtime for parallel
execution.

Lines 19-27 are responsible for partitioning the circuit into {\tt MAX\_PIECES}
pieces that can be worked on in parallel.  In order to partition a region, we
need to provide a {\em coloring}, which is a relation over the valid elements
in a region and a set of colors.  A {\em partition} is an object which given
a coloring and a parent region, will contain a subregion of the parent region
for every color with each subregion containing the elements corresponding 
to the elements colored with the subregions color.

The first step in partitioning the graph
is to partition it into two sets of nodes consisting of 
{\em private} and {\em shared} nodes (lines 20-21).  Private nodes
will only be touched by a single task in a phase, while shared nodes may
be referenced by multiple tasks in a phase.  The thick lines in figure \ref{sfig:part_fig:pvs}
show how the nodes in a small graph might be partitioned into 3 subsets.  Gray
nodes correspond to shared nodes while black nodes are private.  Note that
this partition is disjoint since each node will only be colored once.
%Our current implementation uses small
%integers or booleans for the names of colors, but any enumerable set would be
%fine.  The coloring relation need not be total - it can leave out some elements entirely.  And although it is not required to be injective (it can map a single
%element to multiple colors), the fact that injective mappings will always
%result in disjoint subregions makes them strongly encouraged (when possible).
%Our example uses four colorings to create circuit partitions.

>>>>

In this section we give an informal overview of Legion through an example.
Listing~\ref{lst:code_ex} shows code for an electrical
circuit simulation, which illustrates the core features of the programming model.

The circuit simulation takes a collection of
of wires and nodes where wires meet.  
The simulation repeatedly updates
currents, distributes charges, and updates voltages for as many
time steps as the simulation demands.  We first
describe the partitioning of regions in Section~\ref{subsec:partitioning}
and then describe the data type declarations 
and the constraints they enforce on the simulation that enable
the runtime to efficiently execute the application
in Section~\ref{subsec:datatypes}.

@@

Line 18 declares the main simulator function, which takes a {\tt
Circuit} to be simulated.  This function specifies the regions it 
accesses and the {\em privileges} and {\em coherence} it requires of those regions.
In this case, the {\tt RWE} annotation specifies that the function
accesses the regions {\tt c.r\_all\_nodes} and {\tt c.r\_all\_wires}
with read-write privileges and {\em exclusive} coherence (i.e., no other
function can access the regions concurrently).  The simulation 
reads and writes all nodes and wires, and it must be done
with exclusive access to ensure correct results.  Privileges specify what
the function can do with the regions; coherence specifies what other
functions can do with the regions concurrently.  Functions that
declare their accessed regions, privileges, and coherence are called {\em tasks}
and are the unit of parallel execution in Legion.

Lines 19-27 are responsible for partitioning the circuit into {\tt MAX\_PIECES}
pieces that can be worked on in parallel.  To partition a region, we
provide a {\em coloring}, which is a relation on the elements
of a region and a set of colors.  A {\em partition} is an object which given
a coloring and a region $r$ contains for each color $c$ a {\em subregion} of $r$
with all the elements of $r$ colored $c$.

The region of all nodes in the circuit is first partitioned into 
{\em private} and {\em shared} nodes (lines 20-21).  Private nodes
are touched by a single task in a phase, while shared nodes may
be referenced by multiple tasks in a phase.  The thick lines in figure \ref{sfig:part_fig:pvs}
show how the nodes in a small graph might be partitioned into three subsets.  Gray
nodes are shared and black nodes are private.  Note that
this partition is disjoint since each node has one color.
%Our current implementation uses small
%integers or booleans for the names of colors, but any enumerable set would be
%fine.  The coloring relation need not be total - it can leave out some elements entirely.  And although it is not required to be injective (it can map a single
%element to multiple colors), the fact that injective mappings will always
%result in disjoint subregions makes them strongly encouraged (when possible).
%Our example uses four colorings to create circuit partitions.
